### 2021-08-25 plos_latex_template.tex
RJ: new bib file and some of Andrea's changes to rule 1

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. The larger the conference, the larger the opportunities to network and learn from your cohort. However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arend_disparity_2019, timperley_he_2020, gewin_what_2019, brown_ableism_2018}.

@@

The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperley_he_2020}.

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. There are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built. Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.

@@

To accomplish this diversity vision, you will have to take some conscious decisions toward removing bias from your conference's spaces. Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input \ref{rule_organizing_team}, but also externally because representation matters \ref{rule_communication}.

While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well. Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, are some examples of decisions that can make inclusion real. 
%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}). %and expressing this welcoming spirit in your communication strategy (social networks, website, brand, and visual identity) and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too. 
However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).

@@

\section{Rule X: Have a strong online component of the conference} 
\label{rule_online}
In-person interaction is priceless; however, it is more expensive for some, even unattainable for others. This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arend_disparity_2019, gewin_what_2019}. Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities. This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinson_r_medicine_2021}. Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipour_evaluating_2020, niner_better_2021}; in a study case, conference attendance accounted for $35\%$ of a researcher's footprint \cite{achten_carbon_2013}.

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{niner_better_2021}.

\section{Rule X: Have an inclusive and diverse organizing team}
\label{rule_organizing_team}
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually marginalized groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event). To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. But this will only work if there is real inclusion. One would be surprised by how often conference leads tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in. For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities. The same holds for other dimensions of diversity. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear. Even if creating and maintaining such a team and space is challenging, the positive outcomes far outweigh any minor inconveniences (e.g. time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hong_groups_2004}.

@@

\section{Rule X: Make the conference accessible to all, from start to end}
\label{rule_accessibility}
Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{price_access_2009}. %People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
Inaccessible practices exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

@@

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irish_increasing_2020}. 
Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}).

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaro_your_2016}. The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{aurora_how_2018}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.

@@

\section{Rule X: Search proactively for the best people the conference's vision}
\label{rule_presenters}
When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject. Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyer_notice_2021, sarabipour_evaluating_2020}. Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

@@

On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipour_evaluating_2020, andalib_postdoc_2018, kaplan_postdoc_2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. You can take into account the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arend_disparity_2019}). Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{niner_better_2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier. Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

% % Either type in your references using
% % \begin{thebibliography}{}
% % \bibitem{}
% % Text
% % \end{thebibliography}
% %
% % or
% %
% % Compile your BiBTeX database using our plos2015.bst
\bibliography{references}
% % style file and paste the contents of your .bbl file
% % here. See http://journals.plos.org/plosone/s/latex for 
% % step-by-step instructions.
% % % 
% \begin{thebibliography}{10}

>>>>

Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions. The larger the conference, the larger the opportunities to network and learn from your cohort. However, conferences can become an exclusive space for privileged groups (e.g., white, male, from a rich country, English-native speaker, with no physical disabilities)  \cite{arendDisparityConferenceRegistration2019, timperleyHeMoanaPukepuke2020, gewinWhatScientistsShould2019, brownAbleismAcademiaWhere2018}.

@@

The first step towards a diverse and inclusive conference is to recognize that people are diverse, that some groups face discrimination and might be underrepresented in your community and event \cite{timperleyHeMoanaPukepuke2020}.

When thinking about diversity, each one of us can have different points of view about the meaning of the term, depending on where we come from. If you are reading this, you may be here because you already have 
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity,  geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few. While human diversity should be celebrated and respected in every space, the truth is that there are implicit hierarchies along these axes, and some statuses (e.g., cisgender, white, male, from the US or Europe) hold the privilege of being the default settings for which all systems are consciously and unconsciously built. Thinking about diversity and inclusion is thinking about counteracting the structures that hold these hierarchies in place.

@@

To accomplish this diversity vision, you will have to take some conscious decisions toward removing bias from your conference's spaces. Having diverse people in decision-making positions will affect positively all the other aspects of your conference, not only internally because all the decision-making process will benefit from their input (see Rule \ref{rule_organizing_team}), but also externally because representation matters[see Rule \ref{rule_communication}].

While you may want to improve representation towards some of the most visible dimensions of human diversity, such as race, gender, and country of origin, building a truly inclusive environment means taking care of all the other aspects as well. Having consideration of religious practices, setting specific accommodations for lactating women and child care, having LGBTIQ+-friendly spaces, creating community-only spaces, enforcing the use of pronouns, and treating gender as a non-binary variable, and accomodating are some examples of decisions that can make inclusion real. 
%Representation is an important aspect of diversity and inclusion: seeing people like you, from your country, that speak your language is one of the best way to feel you belong
Some teams gather this vision in a diversity statement (e.g., \url{https://user2021.r-project.org/about/diversity_statement/}) to express the values of the conference and make diversity part of the communication strategy see Rule \ref{rule_communication} and in every decision you make, will let people know that they are seen, respected, and welcome; that this is their space and community too. 
However, it is important to not stop with a statement alone, but to identify specific goals and tasks that ensure compliance with these larger goals, for example, identifying accessible platforms, developing a detailed code of conduct and a team to enforce it, and preparing accessibility guidelines (see other Rules below).

@@

\section{Rule X: Have a strong online component of the conference} 
\label{rule_online}
In-person interaction is priceless; however, it is more expensive for some, even unattainable for others. This inaccessibility is particularly true for global conferences that usually take place in high-income countries, making it financially demanding for international participants and often impossible to attend due to immigration requirements \cite{arendDisparityConferenceRegistration2019,gewinWhatScientistsShould2019}. Online conferences are more inclusive: they do not need a visa or a big budget, and are more accessible to people who may be unable to travel because of health issues or family responsibilities. This means that online conferences have a greater reach, not only in terms of participants but in terms of the tutors and presenters that can participate \cite{atkinsonJournalMedicine20202021}. Furthermore, an online format is more environmentally friendly since it eliminates travel-related emissions \cite{sarabipourChangingScientificMeetings2021,ninerBetterWhomLeveling2021, gattrellComparisonCarbonCosts}.

Alternatively, a conference could have a hybrid format with an in-person and online component. This dual format could allow a group of people to interact face-to-face while providing many others the opportunity to participate remotely. The challenge in this kind of setting would be to make the online component as relevant as the in-person component and not just a consolation prize to the less privileged in the community \cite{ninerBetterWhomLeveling2021}.

\section{Rule X: Have an inclusive and diverse organizing team}
\label{rule_organizing_team}
A genuinely inclusive conference can only be organized by an inclusive and diverse organizing team. Build a team with people from different regions, genders, socioeconomic statuses, and other aspects of diversity. Particular attention should be paid to the usually marginalized groups (see \textbf{Rule 1} discussing the dimensions of diversity lacking in your community and event). To ensure a deeper understanding of the challenges in different diverse groups, it is essential to create a representative working group that functions as a snapshot of the community at large. But this will only work if there is real inclusion. One would be surprised by how often conference leads tend to do everything themselves instead of reaching out to the right people for the right task -- likely because the latter involves careful planning and communication well ahead of time or because they naively overestimate their expertise in areas they lack practical experience in. For example, a cisgender or abled person estimating what might work best for transgender people or people with sensory disabilities. The same holds for other dimensions of diversity. A truly inclusive and welcoming space is one in which everyone in the team is invited and allowed to bring their experience to bear. Even if creating and maintaining such a team and space is challenging, the positive outcomes far outweigh any minor inconveniences (e.g. time zones), and the team will grow, learn, and embrace more inclusive practices \cite{hongGroupsDiverseProblem2004}.

@@

\section{Rule X: Make the conference accessible to all, from start to end}
\label{rule_accessibility}
Conferences are often inaccessible to people with hearing, visual, and other disabilities \cite{priceAccessImaginedConstruction2009}. %People may also find conferences inaccessible due to their caring responsibilities, dietary restrictions, religious practices, or LGBTQ+ status.
Inaccessible practices exclude people from conferences, while accessible practices are inclusive and can be beneficial to a broad spectrum of people. For instance, having captions is helpful to deaf and hard-of-hearing people, non-native speakers, and everyone in general.

@@

Most importantly, accessibility practices are not afterthoughts that can be dealt with at the last minute. 
They require time and early decision-making \cite{irishIncreasingParticipationUsing2020}. 
Conversely, inaccessible decisions are hard to course-correct, e.g., when finding out too late that a venue is inaccessible for wheelchairs in an in-person conference. 
Welcoming people that are usually disabled by standard practices in conferences into the organizing team would allow them to take part in the decisions from the beginning and spot right away the inaccessible practices that need to improve (see \textbf{Rule \ref{rule_organizing_team}}).

@@

To ensure an inclusive, friendly, and safe space in the conference, the organizers need to adopt a code of conduct and set up a team to enforce it \cite{favaroYourScienceConference2016}. The code of conduct is a document meant to keep the community safe and should state clearly the unacceptable behaviors, the spaces of the conference in which it applies, the consequences for engaging in unacceptable behavior, and the way to report violations \cite{auroraHowRespondCode2018}. 
The code of conduct should be displayed prominently in several spaces of the conference to deter people from unacceptable behavior.

@@

\section{Rule X: Search proactively for the best people the conference's vision}
\label{rule_presenters}
When inviting people as keynote speakers, program committee, session chairs and others, we should always look for the best in their subject. Ensuring diversity in each of these roles is a deliberate process -- if there is not much diversity in the first set of names, it needs to be revised to include people from minoritized groups. If these groups are not showing up in the A-list, it is seldom because of their technical acumen -- it is because we, as organizers, are not used to looking in unfamiliar places to bring in the talent. Many of our biased lists are products of the existing systems that have always privileged some groups of people \cite{dwyerNoticeWhoScience2021,sarabipourChangingScientificMeetings2021}. Rather than deter us, this should encourage us to look further to find great people that are not routinely in the spotlight. We need to go beyond our narrow and often limited networks to look for, reach out to, invite, encourage, and onboard these people until there is ample representation across the diversity spectrum and dimensions.

@@

On the other hand, if we are aiming for inclusiveness and representation, the socioeconomic context of participants should be taken into account when determining the registration rates. Usually, there is a higher fee for people from the industry than for academia. A lower fee for non-profit organizations, government employees, or freelancers should also be considered. It is important to include discounts for students as well as for postdocs (early career researchers) to encourage their participation \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}. Since postdoc (or trainee) statuses are not always well-defined in academia and can vary for each country, their payment category should be explicitly defined. You can take into account the cost of living in each country using conversion factors (e.g. from the International Comparison Program report of the World Bank \cite{arendDisparityConferenceRegistration2019}). Consider that, while it is traditional for employers to provide --at least some --conference support in the Global North, this is not the case everywhere.

@@

In international conferences, English is often the official language. Submissions, presentations, tutorials, and workshops are in English. The platforms, the webpage, and official communications are also in English. While English is the primary language in scientific communication and one official language makes it conducive to communicate widely, opening up the conference to other languages could make it less intimidating to people who are not fluent in English \cite{ninerBetterWhomLeveling2021}. Excluding them may potentially lead to missing innovative contributions due to a language barrier. Advertising the conference in several languages and considering having non-English workshops and presentations (with or without captions in English) could help overcome this barrier. For instance, hosting one international day/session per conference might be a great place to start!

@@

% % Either type in your references using
% % \begin{thebibliography}{}
% % \bibitem{}
% % Text
% % \end{thebibliography}
% %
% % or
% %
% % Compile your BiBTeX database using our plos2015.bst
\bibliography{community-science}
% % style file and paste the contents of your .bbl file
% % here. See http://journals.plos.org/plosone/s/latex for 
% % step-by-step instructions.
% % % 
% \begin{thebibliography}{10}
